---
title: "Murray Cod Larval Clusters in the Murrumbidgee River 2013"
author: "Alan Couch"
date: "Thursday, August 14, 2014"
output: html_document
---

In this analysis it is important to remember a few points when interpreting the analysis. Firstly, the larvae come from a nest some unknown distance **above** the site Name, not neccessarily at that site although that is where they were collected. Secondly, it is assumed that the female travelled an unknown distance upstream to deposit her eggs in the nest. It is not unknown however for females to travel downstream to spawn (citation needed). 

It is also important to remember this analysis is a hypothesis generation exercise regarding how many nests are likely represented by the larvae sampled. It is not a post-hoc assignation therefore. The number of nests hypothesesised based on biogeochemistry will need to be validated by genetic analysis.

The movement of the females upstream and the larval dispersal downstream conspire to confound interpretation of the spatial distribution based on biogeochemistry alone.

It should also be remembered that this analysis will ultimately have both elements and C and N isotopes which may improve (or weaken) assignment to clusters.


```{r, "Install Packages and Set Global Options", echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(maps)
library(sparcl)
library(ape)
library(phytools)
library(psych)

opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r,"Import Data and Rename some long variable names"}

qslAllLarvaInfo <- read.csv("C:/Users/s428825/Google Drive/PhD Data Files/Analyses/CNtrial/qslAllLarvaInfo.csv")
# Notice that this data comes from a database query that links three tables - collection, larvae and otolith records.

#Start with base data
df<-data.frame(qslAllLarvaInfo)

#Rename a few long variable names, mean otolith length and distance to AC for example. 
names(df)[names(df) == 'Mean.Otolith.Length.is.in.Millimetres.for.comparison.with.Adults'] <- 'meanOL'
names(df)[names(df) == 'Distance.to.Angle.Crossing..m.'] <- 'distAC'
names(df)[names(df) == 'LarvalRecords.LarvaID'] <- 'LarvaID'
```

#Including DC13, D15N and Mean Otolith Length aka Mummies.

In this case I have standardised the variables iaw Milligan et al 1988, (value/range) which gives more accurate assignation to clusters under a variety of conditions.

```{r, echo=FALSE, "Standardise Data and then Cluster"}
#Start with base data
df<-data.frame(qslAllLarvaInfo)

#Rename a few long variable names, mean otolith length and distance to AC for example. 
names(df)[names(df) == 'Mean.Otolith.Length.is.in.Millimetres.for.comparison.with.Adults'] <- 'meanOL'
names(df)[names(df) == 'Distance.to.Angle.Crossing..m.'] <- 'distAC'
names(df)[names(df) == 'LarvalRecords.LarvaID'] <- 'LarvaID'

# Pare Down to what the Variables needed for clustering
df <- subset(df, YearOnly == 2013, select = c(LarvaID,Site.Name,meanOL,Delta13C,Delta15N))
rownames(df) <- df$LarvaID
df<-subset(df, select = -LarvaID)

# Pare down records to those with delta 13 C values. ie:reduce rows 
df <- subset(df, !is.na(Delta13C))

df$smeanOL<-df$meanOL/(max(df$meanOL) - min(df$meanOL))
df$sDelta13C<-df$Delta13C/(max(df$Delta13C) - min(df$Delta13C))
df$sDelta15N<-df$Delta15N/(max(df$Delta15N) - min(df$Delta15N)) 

df<-subset(df, select = c(Site.Name,smeanOL,sDelta13C,sDelta15N))

#Create Cluster
distxy <- dist(df)
hClusters <- hclust(distxy)

#Basic Plot with 
plot(hClusters,labels=(df$Site.Name), hang = -1, main="Mummy")
#and with larvaID
plot(hClusters,labels=(df$row.names), hang = -1, main="Mummy")

   
# colors the leaves of a dendrogram
y = cutree(hClusters, 12)
ColorDendrogram(hClusters, y = y, labels=df$Site.Name, main = "Mummy", branchlength = 3)

#########
# load code of A2R function
source("http://addictedtor.free.fr/packages/A2R/lastVersion/R/code.R")
# colored dendrogram
op = par(bg = "#EFEFEF")
A2Rplot(hClusters, k = 12, boxes = FALSE, col.up = "gray50")


##Heatmap
set.seed(143)
##Remove non-numerics for matrix construction as Heatmap requires
df <- subset(df, select = c(smeanOL,sDelta13C,sDelta15N))
dataMatrix<-as.matrix(df)
heatmap(dataMatrix)

##MDS
d <- dist(df) # euclidean distances between the rows
fit <- cmdscale(d,eig=TRUE, k=2) # k is the number of dim
#fit # view results

# plot solution 
x <- fit$points[,1]
y <- fit$points[,2]

plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",main="Metric  MDS",  type="n")
text(x, y, labels = row.names(df), cex=.7)
```

In this case there are `r nrow(df)` rows ( `r row.names(df)` ) and `r ncol(df)` columns (`r names(df)`) in the dataframe.

#This Time the Analysis Includes a estimate of Nest Location based on available dispersal days and capture location

This analysis includes an 'arbitrary' nest location calculation based on age of the larva as determined from otolith length, the time it is likely to have had available for dispersal (age - 7 days), and distance above the capture site calculated at 300 metres movement per day. The formula in r is as follows:

```{r, echo =TRUE, eval=FALSE}
#Create proxy Nest site variable called 'nest'.
df$nest<-df$distAC-(df$OLEstimateAge-7)*300
```

##Cluster Analysis Based on Standardised Variables (Milligan and Cooper 1988) 

In this case I have standardised the variables iaw Milligan et al 1988, (value/range) which gives more accurate assignation to clusters under a variety of conditions.

An example of this code is:

- $$df$sDelta13C<-df$Delta13C/(max(df$Delta13C) - min(df$Delta13C))$$

**Needs verification from body of paper (abstract used).**

```{r, echo=FALSE, "Standardise Data and then Cluster with Nest Site Estimate"}
#Start with base data
df<-data.frame(qslAllLarvaInfo)

#Rename a few long variable names, mean otolith length and distance to AC for example. 
names(df)[names(df) == 'Mean.Otolith.Length.is.in.Millimetres.for.comparison.with.Adults'] <- 'meanOL'
names(df)[names(df) == 'Distance.to.Angle.Crossing..m.'] <- 'distAC'
names(df)[names(df) == 'LarvalRecords.LarvaID'] <- 'LarvaID'

# Pare Down to what the Variables needed for clustering
df <- subset(df, YearOnly == 2013, select = c(LarvaID,Site.Name,OLEstimateAge,distAC,Delta13C,Delta15N))

#Create proxtNest site variable.
df$nest<-df$distAC-(df$OLEstimateAge-7)*300

rownames(df) <- df$LarvaID
df<-subset(df, select = -LarvaID)

# Pare down records to those with delta 13 C values. ie:reduce rows 
df <- subset(df, !is.na(Delta13C))

#Standardise according to Milligan

df$snest<-df$nest/(max(df$nest) - min(df$nest))
df$sDelta13C<-df$Delta13C/(max(df$Delta13C) - min(df$Delta13C))
df$sDelta15N<-df$Delta15N/(max(df$Delta15N) - min(df$Delta15N)) 

df<-subset(df, select = c(Site.Name,snest,sDelta13C,sDelta15N))

#df<-subset(df, Site.Name != "Bullen Range")

#Create Cluster
distxy <- dist(df)
hClusters <- hclust(distxy)

#Basic Plot with 
plot(hClusters,labels=(df$Site.Name), hang = -1, main = "Chemistry and Nest Clusters")
#and with larvaID
plot(hClusters,labels=(df$row.names), hang = -1, main = "Chemistry and Nest Clusters")

   
# colors the leaves of a dendrogram
y = cutree(hClusters, 12)
ColorDendrogram(hClusters, y = y, labels=df$Site.Name, main = "Chemistry and Nest Clusters", branchlength = 3)

#########
# load code of A2R function
source("http://addictedtor.free.fr/packages/A2R/lastVersion/R/code.R")
# colored dendrogram
op <- par(bg = "#EFEFEF")
A2Rplot(hClusters, k = 12, boxes = FALSE, col.up = "gray50")

##Heatmap
set.seed(143)
##Remove non-numerics for matrix construction as Heatmap requires
df <- subset(df, select = c(snest,sDelta13C,sDelta15N))
dataMatrix<-as.matrix(df)
heatmap(dataMatrix, main = "Chemistry and Nest Clusters")

##MDS
d <- dist(df) # euclidean distances between the rows
fit <- cmdscale(d,eig=TRUE, k=2) # k is the number of dim
#fit # view results

# plot solution 
x <- fit$points[,1]
y <- fit$points[,2]

plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",main="Chemistry and Nest Clusters MDS",  type="n")
text(x, y, labels = row.names(df), cex=.7)

rm(x); rm(y); rm(d); rm(fit); rm(distxy); rm(op)#cleanup temporary variables
```

In this case there are `r nrow(df)` rows (`r row.names(df)`)  and `r ncol(df)` columns (`r names(df)`) in the dataframe.



